---
title: "Group by and summarize"
output: github_document
---

```{r message = FALSE}
library(tidyverse)

#set the figures' size
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

#set the theme of all graphs
theme_set(theme_minimal() + theme(legend.position = "bottom"))

#set color of all graphs
options(
  ggplot2.continuous.colour = "viridis", #color 边界
  ggplot2.continuous.fill = "viridis" #color 填充
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load the weather dataset
```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode( #add a new variable named `name` and match with the `id`
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>% #Add a new variable to describe the "month" variable rounding the date to the nearest month.
  select(name, id, everything())
```

### `group_by()`
```{r}
weather_df %>% 
  group_by(name, month) %>% #group name and month
  ungroup(month) #ungroup month
```

### Counting things
 - How many observations in ...? 
 
1. `group_by` first, them `summaarize()`
```{r}
#Count month observations:
weather_df %>% 
  group_by(month) %>% #group by month
  summarize(n_obs = n()) #`n()`: how many observations in current group!

weather_df %>% 
  group_by(name, month) %>% #group by name and month
  summarize(n_obs = n()) # how many observations in current group!
```
OR:

2. `count()`
```{r}
weather_df %>% 
  count(name, month, name = "n_obs") #count the name and month (similar as groupby and summarize)
```


### Creating 2x2 tables:
1. 
 - Summarize produces is a data frame, so we can make plots.
 - **Never** use base R's `table` function, since it produces data frame that cannot do anything on.
```{r}
#never use!
weather_df %>% 
  pull(name) %>% 
  table()
```

2. creating 2x2 tables (2ways): (a) and (b) produce the same things! 
```{r}
#a)
weather_df %>% 
  janitor::tabyl(month, name)

#b)
weather_df %>% 
  group_by(name, month) %>% #group by name
  summarize(n_obs = n()) # how many observations in each name and month

weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date)) #show number of distinct values of date in each month.
```

3. 
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when( #IF, THEN function!
      tmax < 5 ~ "cold", # if tmax < 5, then "cold" variable will print "cold"
      tmax >=5 ~ "not_cold",  # if tmax >= 5, then "cold" variable will print "not_cold"
      TRUE      ~ "" #otherwise, ...?
    )
  ) %>% 
  group_by(name,cold) %>% 
  summarize(count = n())

#OR

weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when( #IF, THEN function!
      tmax < 5 ~ "cold", # if tmax < 5, then "cold" variable will print "cold"
      tmax >=5 ~ "not_cold",  # if tmax >= 5, then "cold" variable will print "not_cold"
      TRUE      ~ "" #otherwise, ...?
    )
  ) %>% 
  janitor::tabyl(name, cold)
```

4. Formatting table outputs
```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE) #only show the summarized variables
  ) %>% 
  knitr::kable(digits = 2) #make table with 2 sigfits
```

### General summaries
1. use the produced dataframe to plot ggplot
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE), #drop missing values in tmax variables,
    median_tmax = median(tmax, na.rm = TRUE),
    sd_tmax = sd(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
  geom_point() + 
  geom_line()
```

2. Summarize many columns:
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean)) #mean of variables from prcp to tmin (use across to summarize many columns by using same function)
```

3. Make results easier to read:
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 1) #1 sigfit
```

### `group_by` and `mutated`
 - for example: so we can calculate the mean based on the group (calculate the mean for one group first, and so on)
1. 
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE), #adding a variable for mean_tmax
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

2. ranking function
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(desc(tmax)) #rank tmax largest to lowest; no desc() then from lowest to largest
  ) %>% 
  filter(tmax_rank == 1) #filter the hottest day
```

3. Lagged function

 - "compute day to day temperature variability"
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged_tmax = lag(tmax, n = 1), #show what is the tmax 1 day before (since n = 1), so the table shows the tmax for today and the day before
    tmax_diff = tmax - lagged_tmax
  ) %>% 
  summarize(
    diff_sd = sd(tmax_diff, na.rm = TRUE),
    diff_max = max(tmax_diff, na.rm = TRUE) )
```

### Limitations
what if my "summary" is a linear model...
Unable to do it. 
```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(cor_tmin_tmax = cor(tmin, tmax, use = "complete")) #calculate the correlation btw tmax and tmin


weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  lm(tmax ~ tmin, data = .) #make a linear equation by tmax and tmin, by using this dataset, weather_df

#unable to do 
# weather_df %>% 
  # group_by(name) %>% 
  # summarize(lm = lm(tmax ~ tmin))
```








