---
title: "ggplot1"
output: github_document
---

load in a dataset that we will use often.

```{r}
library(tidyverse)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors( # download NOAA weather data
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


### Scatterplot

1. plot: tmax vs tmin
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```


1) you can save ggplots
```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tmin
```


2. Addcolor? lines? other stuff??
 - transparency of points made by this geom_point() #(aes(color = name)) in  geom_point() but not in ggplot() will
 - se: put confidence interval around the plot 
        (se = FALSE): exclude standard error bar of confidence interval
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3) + 
  geom_smooth(se = FALSE)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) + #draw best fit line and exclude standard error bar of confidence interval
  facet_grid(. ~ name)
```


Make more scatterplots
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, size = prcp)) +
  geom_point(alpha = .3) +
  facet_grid(. ~ name) +
  geom_smooth(se = FALSE)
```


### Use data manipulation as a part of this
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```


### Stacking geoms

which geoms do you want?

1. Draw smooth curve without scatterpoints 
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth()
```


2. geom_hex(): give density of data
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex()
```



