---
title: "visualization_ggplot2"
output: github_document
---

load in a dataset that we will use often.

```{r message = FALSE}
library(tidyverse)
# library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(             #download NOAA weather data, pull data names c, var, and date
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


### Scatterplot:

1. scatterplot: tmax vs tmin
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

# or

ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```


1) save the ggplots and can print later
```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tmin

#or

ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

ggp_tmax_tmin +  geom_point()
```


2. Add color/ smooth curves/ other stuff
 - se: put confidence interval around the plot 
        (se = FALSE): exclude standard error bar of confidence interval
 - alpha = .3: transparency of color of the points
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + #color all ggplots based on names 
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) + #draw the smooth curves in the middle of dataset, and exclude standard error bar of confidence interval
  facet_grid(. ~ name) #multi-panel graphs: separate the plot based on variable `name`

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3) +  #only color the scatterplot based on names 
  geom_smooth(se = FALSE) #since only color the scatterplot, only show one smooth curve based on entire data
```


1) Make more scatterplots
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) + #let the size of plots based on prcp's value (larger prcp value, larger point)
  geom_point(alpha = .3) +
  facet_grid(. ~ name) +
  geom_smooth(se = FALSE, size = 0.8)
```


### Use data manipulation as a part of this
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```


### Stacking geoms

which geoms do you want?

1. Draw smooth curve without scatterplot
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth()
```


2. give density of data (neat geom!)
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex()

weather_df %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_density2d() +
  geom_point(alpha = .3)

    
```


### Univariate plots

# Everything before ggplot use pipe, %>%; Everything after ggplot use plus, +.

1. Histogram
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + # fill = name, fill the color of histogram bars based on diff names.
  geom_histogram() +
  facet_grid(. ~ name) # separate one plot to separate plots based on names (.~ nmaes: 1 rows, multiple columns)

```

2. Density plot: smoothed down the histogram (easier to compare, but loose some data)
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3) #color's transparency of density plots is 0.3
```

3. Boxplot
```{r}
weather_df %>% 
  ggplot(aes(y = tmax)) + #one boxplots of tmax values
  geom_boxplot()


weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) + #draw multiple boxplots based on names
  geom_boxplot(alpha = 0.3)
```

4. Trendy plot: violin plots :-)
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.3) +
  stat_summary(fun = "median") #give the point of median
```

5. Ridge plots
```{r}
library(ggridges)

weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(alpha = .8, scale = .8) #multiple density curves based on different names
```

* color = name vs. fill = name:    
 color = name: color the outline    
 fill = name: fill the color in the graph   

### Save and Embed plots

1. Save a scatterplot
```{r}
weather_plot = 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + #aes(color = name), since it is aesthetic change.
  geom_point(alpha = .3)

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5) #save the plot `weather_plot` as the name "weather_plot.pdf", with specific size
```
2. Embedding
```{r, fig.width = 12, fig.height = 7}
#make the figure 12x7

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)
```



