---
title: "visualization_ggplot2"
subtitle: "Part_II"
output: github_document
---

```{r message = FALSE}
library(tidyverse)
#library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors( # download NOAA weather data
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


### Labels of ggplot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from the rnoaa package with three stations, temperatures in 2017." #说明
  )

```


### Scales
1. data scales
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from the rnoaa package with three stations"
  ) +
  scale_x_continuous(         #scale_x_continuous (since x is continuous variable)
    breaks = c(-15, 0, 15),   #scale the x variables into 3 breaks    
    labels = c("-15C", "0", "15C") #label the scales of x
  )  +
  scale_y_continuous(
    trans = "sqrt",         #transformation the y data to square root
    position = "right"      #change the scale to the right of plot
  )
```


2. color scales
```{r}
library(viridis) #for viridis
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from the rnoaa package with three stations"
  ) +
   scale_color_hue(
     name = "Location", #change the name of caption on the rightside
     h = c(100, 300)) #手动设色系
  
  
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from the rnoaa package with three stations"
  ) +
   viridis::scale_color_viridis(
     name = "Location", #change the name of caption on the rightside
     discrete = TRUE
   )
```

### Themes
1. shift the legend
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from the rnoaa package with three stations"
  ) +
  scale_color_viridis_d() +
  theme(legend.position = "bottom") #change the legend to the bottom
#change position should be the last step
```

2. Change the overall theme
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from the rnoaa package with three stations"
  ) +
  scale_color_viridis_d() +
  theme_bw()   #black outline, white background

#Prefer 这个！
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from the rnoaa package with three stations"
  ) +
  scale_color_viridis_d() +
  theme_minimal() #get rid of the black outline of graph
```


# copy-and-pasted to the beginning of every rMarkdown document:
**code for setting options of graphs that will use**
```{r eval=FALSE}
library(tidyverse)

#set the figures' size
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

#set the theme of all graphs
theme_set(theme_minimal() + theme(legend.position = "bottom"))

#set color of all graphs
options(
  ggplot2.continuous.colour = "viridis", #color 边界
  ggplot2.continuous.fill = "viridis" #color 填充
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### combind two different datasets: two whole datasets are in different graph types
 - `data` args in `geom`

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

waikiki %>% 
ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)

#or: 

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

### Put up different types plots side by side:
 - `patchwork`
```{r}
library(patchwork)

#three plots
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none") #delete legend

ggp_prcp_dens = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = 0.5)

ggp_tmax_date = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth()

#combine
(ggp_tmax_tmin + ggp_prcp_dens + ggp_tmax_date) # combining different ggplots on the same row
(ggp_tmax_tmin + ggp_prcp_dens) / ggp_tmax_date # two plots on the top, one on the bottom
  
```


### data manipulation
1. reorder the graph should reorder the data:
```{r}
#plot by the x = name with ascending tmax:

##Should reorder the dataset first, but not change in ggplot

weather_df %>% 
  mutate(
    name = factor(name), #change the name from character to factor
    name = forcats::fct_relevel(name, c("Waikiki_HA")) #reorder the name, let "wakiki" be first
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +  
  geom_boxplot(alpha = .5)


weather_df %>% 
  mutate(
    name = fct_reorder(name, tmax)  #change the name from character to factor, Also reorder the name by ascending tmax.
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +  
  geom_boxplot(alpha = .5)
```

2.Want two variables in a graph
 - what if I want densities for tmax and tmin simultaneously?
```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observations",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observations)) + 
  geom_density(alpha = 0.3) +
  facet_grid(. ~ name)
```


### Revisit the pups
 - combine data manipulations and ggplot 
 
1. Example1: 
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_data = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

#left join:
fax_data = left_join(pup_data, litters_data, by = "litter_number")


#Draw graphs between two variables in this huge datasets
fax_data %>% 
  ggplot(aes(x = dose, y = pd_ears)) +
  geom_violin() +
  facet_grid(.~ day_of_tx)


#What if I am interested in the relationship between two variables from 4 types (which are begining with "pd_")
##combine the 4 variables first
fax_data %>% 
  select(dose, day_of_tx, starts_with("pd_")) %>%  #select the data I am interested in: the values of two variables that are cooresponding to 4 types
  pivot_longer(      #reorganized the data that I am interested in into two variables
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day"
  ) %>% 
  drop_na() %>% #drop missing values
  mutate(outcome = forcats::fct_relevel(outcome, "pd_ears","pd_pivot","pd_walk","pd_eyes")) %>% 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(day_of_tx ~ outcome)

```

2. Example2:

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  mutate(visit = recode(visit, "bl" = "00m"))

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_point() +
  geom_line(aes(group = id)) #good for longitude data analysis. # one person observe for a period of time
```


















